---
title: "Assignment 2 Starter Analysis"
format: html
---

## Setup

```{r}
#| message: false
#| warning: false
library(tidyverse)  # AI-EDIT(2026-06-23): tidyverse-default — consolidated core library() calls into library(tidyverse)
```

## Load Data

```{r}
nhanes <- read_csv("examples/nhanes-equity/data/nhanes_equity_v6.csv", show_col_types = FALSE)
```

## Inspect

```{r}
dim(nhanes)
names(nhanes)
glimpse(nhanes)
colSums(is.na(nhanes))
```

## Summarize

Replace this example with your own summary if you choose a different question.

```{r}
Age_summary <- nhanes |>
  filter(!is.na(Age), !is.na(Cycle)) |>
  group_by(Cycle) |>
  summarise(
    n = n(),
    mean_Age = mean(Age, na.rm = TRUE),
    .groups = "drop"
  )

Age_summary
```

## Function

```{r}
summarize_column <- function(data, column_name) {
  if (!column_name %in% names(data)) {
    stop("Column not found: ", column_name)
  }

  values <- data[[column_name]]
  if (!is.numeric(values)) {
    stop("Column must be numeric: ", column_name)
  }

  observed <- values[!is.na(values)]
  tibble(
    count = length(observed),
    missing = sum(is.na(values)),
    mean = if (length(observed) > 0) mean(observed) else NA_real_,
    sd = if (length(observed) > 1) sd(observed) else NA_real_
  )
}

summarize_column(nhanes, "BMI")
```

## Save Output

```{r}
dir.create("assignments/assignment02-r-with-ai/submission/output", recursive = TRUE, showWarnings = FALSE)
write_csv(Age_summary, "assignments/assignment02-r-with-ai/submission/output/summary_table.csv")
```

## AI Audit Note

## Audit trail
AI was used to generate Q&A responses and code suggestions. For example How to use dplyr to use the summarise, filter and mutate functions to analyze the NHANES dataset.

## What AI helped with
AI helped clear up code use and their functions. It a

## What I changed
<!-- 1-3 sentences: what you edited, fixed, or rejected. -->

## How I verified the result
<!-- 1-3 sentences: what you checked it against (data, docs, course notes). -->
